% !TEX root = ../main.tex

The final pipeline refits all four models on the full training set (no validation or holdout removed) with the cached per-seed best iterations from the val-tuned models: LightGBM seeds 42, 123, 456 with 423, 316, and 398 trees respectively, and XGBoost with 694 boosting rounds. Predictions on the 4{,}959{,}183-row test set are blended at $w_{\text{xgb}} = 0.5$ and sorted within each \texttt{srch\_id} by descending blended score. The output is written to \texttt{results/submission.csv} with columns \texttt{srch\_id, prop\_id}; smoke tests confirm zero duplicate (search, property) pairs and full coverage of the 199{,}549 test searches.

The Kaggle public NDCG@5 of this submission is 0.41706, with a corresponding locked-holdout NDCG@5 of 0.41521. The Kaggle public score is +0.00185 higher than the holdout, which is opposite to the typical small shrinkage one would project from a peeked-at holdout. The surprise direction is consistent with two stabilising effects we built into the pipeline. First, the locked 80/10/10 group-aware split is enforced across every notebook, so the holdout is never used in Optuna tuning, blend-weight selection, or rerank-$\lambda$ selection. Second, the L2-heavy Optuna configuration reduces \texttt{prop\_id} memorisation, which is the typical source of val-to-public shrinkage on this dataset. The holdout was a slightly pessimistic estimate, not a coincidentally optimistic one. The 0.41706 result is comparable to the Bellucci~\cite{bellucci2021dmt} reference of 0.41726 on the same in-class leaderboard.
