\documentclass[a4paper,10pt]{article}
\usepackage[margin=2cm]{geometry}
\usepackage[utf8]{inputenc}
\usepackage{booktabs}
\usepackage{enumitem}

\title{Process Report: VU-DM-2026-Group-5}
\author{Hendrico Burger \quad Hidde Franke \quad Joanne Pals}
\date{}

\begin{document}
\maketitle
\thispagestyle{empty}

\section*{Workflow and discipline}

We worked through nine model-development stages, beginning with a
bootstrap and EDA stage and progressing through anchor modelling,
feature-hypothesis testing, model comparison and bias mitigation,
ensemble extensions, the historical-prior model, and a final
exploratory stage. Each stage changed one main element on the main
branch, never two, and was preceded by a short written plan that
listed scope, hypotheses, expected impact, and an acceptance
criterion. Acceptance was decided by same-seed deltas on a locked
holdout (sha256 prefix \texttt{bc8ea6f6}, 19{,}980 \texttt{srch\_id}
queries, group-disjoint from train and validation), not by raw means
against a moving anchor. Train-only columns were dropped before
features were built; target encoding used 5-fold group K-fold on
\texttt{srch\_id}, never train-test joint statistics. We pinned
thread counts (\texttt{num\_threads=6}) and never used \texttt{-1}.
Memory peaked under 10\,GB during training; the working laptop was
usable throughout.

\section*{Timeline}

\begin{itemize}[leftmargin=*]
\item \textbf{Bootstrap and EDA stage (2026-05-06).} EDA on full
    data plus a five percent stratified sample, locked holdout JSON
    with sha256 \texttt{bc8ea6f6}, four heuristic baselines on the
    locked holdout (random 0.1579, star 0.1949, price 0.1987,
    location score 2 0.2836).
\item \textbf{LambdaMART anchor stage (2026-05-06).} LightGBM
    LambdaMART on the 49 raw-feature anchor, three seeds, mean
    holdout NDCG@5 = 0.392667 with std 0.001070, three-seed
    score-average 0.395289.
\item \textbf{Feature-hypothesis stage (2026-05-06).} Five
    hypothesis blocks: per-search normalisation features,
    leave-one-out property priors, a mean-position proxy, negative
    downsampling, and competitor and price-composite features. All
    five rejected on same-seed delta against the anchor.
\item \textbf{Model comparison and bias stage
    (2026-05-06 to 2026-05-07).} Hyperparameter exploration
    (label-gain variants, depth, learning rate) rejected;
    documentation check confirmed that
    \texttt{lambdarank\_unbiased} is not a parameter in
    LightGBM 4.x. XGBoost \texttt{rank:ndcg} compared as a
    sample-only second technique (LightGBM 0.329601 versus
    XGBoost 0.356409 on the 5\,\% sample). Bias detection by
    visitor country and by property popularity tertile;
    post-processing exposure rerank with $\lambda = 0.20$
    (global NDCG@5 drop 0.003090; medium and low tertile exposure
    rise by 5.3 and 2.2 percentage points respectively).
    Safe-baseline checkpoint and Kaggle upload
    (\texttt{submit\_anchor\_3seed\_avg.csv}, public 0.39500),
    version-control checkpoint
    \texttt{final\_safe\_baseline\_public\_039500}.
\item \textbf{Ensemble extension stage (2026-05-07).} Five-seed
    extension (seeds 42, 123, 456, 789, 2026) holdout 0.397348;
    with a post-hoc additive blend over per-search z-scored
    signals holdout 0.398246. Kept as a local result, not
    uploaded.
\item \textbf{Diverse-ensemble stage (2026-05-07).} Diverse-config
    LightGBM (depth 8, 63 leaves, learning rate 0.035,
    $\lambda_2 = 10$, bag 0.85, feature 0.75) over 3 seeds,
    weighted 0.60 anchor plus 0.40 diverse with the same blend,
    holdout 0.399638. Manual Kaggle upload, public 0.39685.
\item \textbf{Historical-prior stage (2026-05-07).} Twelve
    leakage-safe historical priors over \texttt{prop\_id},
    \texttt{srch\_destination\_id}, and the pair, with K-fold
    encoding. Three-seed score-average holdout 0.400682 (first
    local result above 0.4). Manual Kaggle upload, public
    0.39690 (final selected submission).
\item \textbf{Final exploratory stage (2026-05-07).} A
    pre-specified rank blend over weight sets W0--W6 reached
    holdout 0.401482 at its maximum but we did not select it; the
    conservative pick at 0.400453 was also not uploaded. A fourth
    prior block keyed by
    \texttt{(prop\_id, visitor\_location\_country\_id)} gave a
    same-seed delta of $+0.000745$, below the $+0.0010$ extension
    threshold, and was rejected.
\item \textbf{Finalisation stage (2026-05-07).} Clean folder
    \texttt{Assignment~2/} produced from the working folder:
    structure plus notebook outlines; figure generation from
    cached parquet and earlier report card JSON; scientific and
    process report drafts; verified Bellucci 2021 citation; final
    read-through and submission package.
\end{itemize}

\section*{Critical reflection}

The most important lesson is the local-to-public shrinkage we
observed: the public-versus-local gap moved from $+0.000289$
(safe baseline) to $-0.002788$ (diverse ensemble) to $-0.003782$
(final submission). Each individual decision was leakage-safe, but
the cumulative search budget on a fixed locked holdout inflated
expected public gains. The conservative choice in the final
exploratory stage (not uploading the holdout-max blend, keeping the
simpler historical-prior 3-seed model as the headline) is our
explicit response to this: we treated holdout NDCG@5 as a selection
signal that loses calibration over many stages and preferred the
candidate whose simplicity reduced the reasoning chain.

A second lesson is the importance of feature provenance: every
feature in the final pipeline traces back to a primary source or
to a verified leakage-safety argument, and each rejected feature
has an experiment record documenting the same-seed delta. This
made it cheap to discard hypotheses without lingering doubt.

\section*{Tooling and process}

\textbf{Documentation.} Short Markdown files in the working folder
track state across sessions: current status, prioritised work,
accept/reject decisions, durable lessons, pitfalls, report-worthy
claims, and a session-by-session timeline that this report draws
on. Each stage also appends a row to a structured CSV log of
experiments, and each public leaderboard upload is recorded on the
same row.

\textbf{Version-control checkpoints.} Each accepted milestone is
captured by a git commit and tag, beginning with a checkpoint at
the safe baseline submission, followed by a local-only checkpoint
after the five-seed blend, a checkpoint after the diverse ensemble
submission, and a checkpoint after the historical-prior submission.
We preserved versioned outputs (the three uploaded submission
CSVs and the relevant per-stage JSON summaries) by sha256 between
sessions and re-verified them at the start and end of each working
session.

\textbf{Leaderboard usage.} We treated the public leaderboard
strictly as an external check, not as a tuning signal. We uploaded
three submissions in total over the project (safe baseline,
diverse ensemble, historical-prior submission). The final
exploratory stage produced two candidate CSVs that we deliberately
did not upload, both because the holdout-max selection was the
wrong move under our shrinkage observation and because
public-leaderboard tuning was outside our chosen protocol.

\section*{Task division}

The work was a collaborative effort and most steps were
cross-reviewed. The lead-contributor split below reflects who
drove each thread day to day.

\begin{itemize}[leftmargin=*]
\item \textbf{Hidde Franke}: project orchestration, the staged
    development plan, the LightGBM modelling track from the
    LambdaMART anchor stage through the historical-prior stage,
    the leakage-safe prior encoding, ensembling, the three Kaggle
    uploads, the clean folder and \texttt{src/} pipeline, and
    report assembly.
\item \textbf{Hendrico Burger}: EDA support and figure inspection,
    feature-engineering review and ablation sanity checks within
    the feature-hypothesis stage, validation of same-seed deltas
    and the locked holdout, and process documentation across the
    development stages.
\item \textbf{Joanne Pals}: the bias and deployment thread within
    the model comparison and bias stage, fairness interpretation
    including the popularity-tertile rerank with $\lambda = 0.20$,
    scientific and process report editing, and process report
    review.
\end{itemize}

\section*{Reproducibility}

The clean folder \texttt{Assignment~2/} contains the deliverable
code under \texttt{src/}, the explanatory notebooks under
\texttt{notebooks/}, and the final submission CSV in
\texttt{results/}. The locked holdout JSON, its sha256 hash, and
the per-seed best-iteration values are in \texttt{artifacts/}. The
command \texttt{python src/pipeline.py --train --predict}
reproduces the final submission from the processed parquet files;
thread counts are pinned and we never use $-1$.
\texttt{python src/pipeline.py --validate
results/submit\_iter07\_prior\_3seed.csv} validates the included
submission CSV without retraining.

\end{document}
